\section{Quick Summary}
\label{sec:started}

Here's a super brief description of how to use Bertini\_real.


\begin{enumerate}
\item Open a terminal or command prompt. Navigate to the directory containing the {\tt input} file.
\item Run Bertini on an input file using the \texttt{tracktype:1} setting. This is done by typing in the command line: \texttt{bertini} with an input file named \texttt{input}. Bertini will produce a Numerical Irreducible Decomposition that will be used by Bertini\_real.
\item Run Bertini\_real on the same input file. Similarly, just type \texttt{bertini\_real} in the command line. Bertini\_real will provide a cellular decomposition of the real portion of a one- or two- dimensional complex algebraic set.
\item Consider running the {\tt sampler} to smooth your decomposition.
\item Visualize the results of Bertini\_real in Matlab. Enter Matlab and call \texttt{gather\_br\_samples}, which parses the output results of  into a .mat file, and then call \texttt{bertini\_real\_plotter}, which will plot the curve or surface in Matlab. 




\end{enumerate}

\subsection*{Notes}

 Please note that 
	\begin{itemize}
	\item either the Matlab executable must be on the path, with the symbolic toolbox;
	\item or, you have Python installed with supporting modules as necessary.  I don't remember what those modules are right now.  {\tt pip install} is your friend.  Decompose Whitney and install dependencies til it works ;)
	\end{itemize}

Bertini\_real can be 
\begin{itemize}
\item downloaded from \url{http://bertinireal.com/download.html}, or 
\item cloned from \url{https://github.com/ofloveandhate/bertini_real}.  
\end{itemize}
Use of Bertini\_real depends on Bertini, which itself has several important dependencies (see Section~\ref{sec:installation}).
Once installed, you can run Bertini\_real from the command line. 


