%
%documentstyle and options : ARTICLE STYLE
\documentclass{acmtrans2m}
%
\usepackage{epsfig,graphicx}
%
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{tabularx}
\usepackage{multicol,lscape}
%\usepackage{rotating}
%
%
% definitions
%
%
%
\usepackage{float}
%\floatstyle{boxed}
%\newfloat{algorithm}{thp}{loa}
%\floatname{algorithm}{Algorithm}
%
%\input macro.tex 
%\input macro_box.tex 
%\input environ_article.tex 
\newcommand{\neuc}[1]{ \left\Vert {#1} \right\Vert_2}
\newcommand{\md}[1]{   \left\vert {#1} \right\vert}
\newcommand{\ds}[2]{\displaystyle{\frac{#1}{#2}}}
\newcommand{\ignore}[1]{}
\newcommand{\trpa}[1]{}
\newcommand{\toms}[1]{#1}

%
%
%\title{An implementation of GMRES \\ for complex linear systems}
\markboth{Val\'erie Frayss\'e et al.}{API of the FGMRES Routines for Real and Complex Arithmetics}

\title{API of a set of FGMRES Routines for Real and Complex Arithmetics on High Performance Computers\footnote{The major part of this work was carried out while the three authors
     were working at CERFACS}}
\author{VAL\'ERIE FRAYSS\'E\\{Kvasar Technology LLC, Boston, USA} \\
LUC GIRAUD\\{ENSEEIHT-IRIT, Toulouse, France}\\
SERGE GRATTON\\{CNES, Toulouse, France}
}
%\date{CERFACS Technical Report TR/PA/97/49 - \today}
\date{}
\begin{document}
\maketitle

%
%
\section{Framework}

The Generalized Minimum RESidual (GMRES) method was proposed by Saad
and Schultz in 1986~\cite{SAA.SCH.86} in order  to solve
large, sparse and non Hermitian linear systems.
GMRES belongs to the class of Krylov based iterative methods.

The FGMRES (Flexible Generalized Minimum Residual) method~\cite{SAA.93} is among the
most widely used Krylov solvers for the iterative solution of general large
linear systems when variable preconditioning is considered.


For the sake of generality we  describe this method for linear
systems that are complex, everything also specialises to real
arithmetic calculation.
Let $A$ be a square nonsingular $n \times n$ complex matrix, and $b$ be
a complex vector of length $n$, defining the linear system
\begin{equation}
A x = b
\label{eq:linsys}
\end{equation}
to be solved.
Let $x_0 \in \mathbb{C}^n$ be an initial guess for this linear system and
$r_0 = b - Ax_0$ be its corresponding residual.

  The convergence of GMRES or GMRES(m) to solve~\eqref{eq:linsys} might be slow.
To overcome this drawback, one often prefers to solve
a transformed linear system that is referred to as the
preconditioned linear system. 
More precisely if $A \approx M^{-1}$ we
actually solve the linear system
\begin{equation}
A M z =  b
\label{eq:preclinsys} 
\end{equation}
with $x = M z$.
In some situation, it might not be possible to explicitly
form $M$ and its application to a vector might vary from
one iteration to the next (for instance $M t$ is obtained
by iteratively solving the linear system $A y = t$).
 Such a preconditioninng strategy is referred to as flexible 
preconditioning.
The FGMRES (Flexible Generalized Minimum Residual) method~\cite{SAA.93} is among the
most widely used Krylov solvers for the iterative solution of general large
linear systems when variable/flexible preconditioning is considered.

\input{implementation}
%
%
\bibliographystyle{acmtrans}
%\bibliography{bib}

\begin{thebibliography}{}

\bibitem[\protect\citeauthoryear{Frayss\'e, Giraud, Gratton, and
  Langou}{Frayss\'e et~al\mbox{.}}{2003}]{fggl:03}
{\sc Frayss\'e, V.}, {\sc Giraud, L.}, {\sc Gratton, S.}, {\sc and} {\sc
  Langou, J.} 2003.
\newblock A set of {GMRES} routines for real and complex arithmetics on high
  performance computers.
\newblock {T}echnical {R}eport TR/PA/03/03, CERFACS, Toulouse, France.
\newblock Available on http://www.cerfacs.fr/algor.

\bibitem[\protect\citeauthoryear{Frayss\'e, Giraud, Gratton, and
  Langou}{Frayss\'e et~al\mbox{.}}{2005}]{fggl:05}
{\sc Frayss\'e, V.}, {\sc Giraud, L.}, {\sc Gratton, S.}, {\sc and} {\sc
  Langou, J.} 2005.
\newblock Algorithm 842: A set of {GMRES} routines for real and complex
  arithmetics on high performance computers.
\newblock {\em ACM Trans. Math. Softw.\/}~{\em 31,\/}~2, 228--238.

\bibitem[\protect\citeauthoryear{Saad}{Saad}{1993}]{SAA.93}
{\sc Saad, Y.} 1993.
\newblock A flexible inner-outer preconditioned {GMRES} algorithm.
\newblock {\em SIAM J. Sci. Comput.\/}~{\em 14}, 461--469.

\bibitem[\protect\citeauthoryear{Saad and Schultz}{Saad and
  Schultz}{1986}]{SAA.SCH.86}
{\sc Saad, Y.} {\sc and} {\sc Schultz, M.} 1986.
\newblock {GMRES}: A generalized minimal residual algorithm for solving
  nonsymmetric linear systems.
\newblock {\em SIAM J. Sci. Stat. Comput.\/}~{\em 7}, 856--869.

\end{thebibliography}
%
\end{document}
