
\subsection{Instructions for OSX}


\subsubsection{Preliminary Work}
\begin{itemize}
  \item Compilers \&c: Install XCode 8 from the App Store.  Open it at least once.  In terminal, type {\tt xcode-select --install} (you must open XCode at least once before doing this)
   
  \item Homebrew: It is highly recommended that Mac users install the program \href{http://brew.sh}{Homebrew} to use to install these packages. Once that has been done, installing the previously listed dependencies becomes simple. In terminal,  type \texttt{brew search \_\_\_\_} to list packages related to \_\_\_\_, where \_\_\_\_ is your search (for example, GMP, Boost, or MPI). To download new software via Homebrew, type in terminal \texttt{brew install \_\_\_\_} 
   \item Matlab: make sure to have it installed, and on the path for the command line.
   \item Python: use {\tt pip} to install `dem dependencies
\end{itemize}


\subsubsection{Installation}
\begin{itemize}
  \item Install parallel Bertini from source
   \begin{enumerate}
    \item Download the source tarball ({\tt .tar.gz}) from \href{http://bertini.nd.edu}{bertini.nd.edu}
    \item Unpack the tarball (just double click it, if you're into mousing)
    \item If you didn't already, now use Homebrew to install the dependencies for Bertini1 \\{\tt brew install autoconf automake libtool mpfr gmp boost mpich}\\ ok, Bertini1 doesn't depend on Boost, but Bertini\_real does
    \item In the terminal, move to the unpacked tarball for Bertini1
    \item {\tt ./configure}
    \item {\tt make -j 4 \&\& make install}
  \end{enumerate}
  \item Install Bertini\_real  
   \begin{enumerate}
    \item {\tt which git}  if it's emtpy, then {\tt brew install git}
    \item {\tt mkdir code \&\& cd code}
    \item {\tt git clone https://github.com/ofloveandhate/bertini\_real}
    \item {\tt cd bertini\_real}
    \item {\tt autoreconf -i}
    \item {\tt ./configure}
    \item {\tt make -j 4 \&\& make install} 
   \end{enumerate} 
\end{itemize} 


Problems?  Raise an issue on GitHub, please. 