% !TeX encoding = UTF-8 Unicode
% !TeX root = manual.tex
% !TeX spellcheck = en_GB

\chapter{Changelog}

\section*{v1.0.7.2}
\begin{itemize}
\item Code changed to Python indentation style
\item version: v1.0.7.1: No error messages if Gurobi is missing
\item Minor changes due to Version R2019a.
\item Rewrote the test-suites. They now are in the standard Matlab script-based unit tests format, but need the \texttt{TTEST}-packages (also included in the package now) in order to work.
\item Fixed a bug in \verb|leadingeigenvector|.
\end{itemize}


\section*{v1.0.6}
\begin{itemize}
\item Made all warning texts to be Matlab warnings. The warning-id's are subject to be changed.
\item Rewrote \verb|findsmp|, making it faster and added the possibility to search for \emph{spectral minimizing products}.
Old implementation is renamed to \verb|findsmpold|.
\item Added options \verb|expect| and $\verb|expecte|$ to check parameters parsed by \verb|parsem|
\item Behaviour change in \verb|removezero|. Row vectors are now considered to be 2-dimensional arrays.
\item Colour options handling of \verb|plotm| improved.
\item Function \verb|tjsr_domain| is nearly finished. Its function name will be changed in a future release.
\item Added using of hyper-spherical coordinates to \verb|cart2sphm|
\item Bugfix in type \verb|3| of \verb|leadingeigenvector|
\item Behaviour change of option \verb|cycle| of \verb|leadingeigenvector|
\item Behaviour change of function \verb|tgenNecklaces|. 
Function now returns \emph{short} $n$-bead necklaces with $k$ colours instead
$n$-bead necklaces with $k$ colours instead.

\end{itemize}

\section*{v1.0.5}
\begin{itemize}
\item Added options \verb|noprint|, \verb|once|, \verb|save| and \verb|load| to \verb|vprintf|.
\item Added options \verb|link| and \verb|equal| to \verb|plotm|
\item Added type \verb|3| to \verb|leadingeigenvector|
\end{itemize}