\begin{thebibliography}{10}

\bibitem{BLAIR93}
J. R. S. Blair and B. Peyton, 
\newblock ``An introduction to chordal graphs and clique trees'', 
\newblock in {\em Graph Theory and Sparse Matrix Computation}, 
A. George, J. R. Gilbert and J. W. H. Liu, eds., 
Springer-Verlag, New York, 1993, pp. 1 -- 29. 

% cited
\bibitem{GAMS}
GAMS HomePage, 
\newblock \verb+http://www.gams.com/+

% cited
\bibitem{GLOBAL}
GLOBAL Library,
\newblock \verb+http://www.gamsworld.org/global/globallib.htm+


\bibitem{KIM03} S.~Kim, M.~Kojima and H.~Waki, ``Generalized Lagrangian duals
and sums of square relaxation of sparse polynomial optimization problems'', 
%B-395, Dept. of Mathematical and Computing Sciences, Tokyo Institute of Technology,
%Oh-Okayama, Meguro, Tokyo 152-8552 Japan, 
{\em SIAM J. Optimization} {\bf 15} (2005) 697--719.

% cited
\bibitem{KOJ03a} 
M.~Kojima, S.~Kim and H.~Waki,
\newblock ``Sparsity in sums of squares of polynomials'', 
\newblock  {\em Mathematical Programming} {\bf 103} (2005) 45-62.

%\bibitem{KOJ03} M.~Kojima, ``Sums of squares relaxations of polynomial semidefinite 
%programs'', 
%\newblock B-397, Dept. of Mathematical and Computing Sciences, Tokyo Institute of 
%Technology, Oh-Okayama, Meguro, Tokyo 152-8552 Japan, November 2003. 

%\bibitem{KOJ02} M.~Kojima, S.~Kim and H.~Waki,
%\newblock ``A general framework for convex relaxation of polynomial
%optimization problems over cones'',
%\newblock {\em Journal of Operations Research Society of Japan}, {\bf 46}  2 (2003) 125-1
%44.

%\bibitem{KOJ04} M.~Kojima and M.~Muramatsu, ``An extension of sums of squares relaxations 
%to polynomial optimization problems over symmetric cones'', 
%B-406, Dept. of Mathematical 
%and Computing Sciences, Tokyo Institute of 
%Technology, Oh-Okayama, Meguro, Tokyo 152-8552 Japan, April 2004. 

% cited
\bibitem{LAS01}
J.~B.~Lasserre:
\newblock Global optimization with polynomials and the problems of
        moments, 
\newblock {\em SIAM Journal on Optimization}  {\bf 11} (2001) 796--817.

\bibitem{LAS06} 
J.~B.~Lasserre:
\newblock Convergent SDP-relaxations in polynomial optimization with sparsity, to appear in 
\newblock {\em SIAM Journal on Optimization}.

% cited
%\bibitem{PARRILO03}
%P. A. Parrilo,
%\newblock ``Semidefinite programming relaxations for semialgebraic problems'',
%\newblock {\em Mathematical Programming}  {\bf 96} (2003) 293-320.

% cited 
\bibitem{STRUM99}
J. F. Strum,
\newblock ``SeDuMi 1.02, a MATLAB toolbox for optimization over symmetric cones'',
\newblock {em  Optimization Methods and Software}, {\bf 11 \& 12} (1999)  625-653.

%cited
\bibitem{SDPT3}
\newblock R. H Tutuncu, K. C. Toh, and M. J. Todd, 
``Solving semidefinite-quadratic-linear programs using SDPT3'', 
{\em Mathematical Programming} {\bf 95}  (2003) 189--217. 

% cited
\bibitem{YAMASHITA03}
M. Yamashita, K. Fujisawa and M. Kojima,
\newblock ``Implementation and evaluation of SDPA 6.0 (SemiDefinite
Programming Algorithm 6.0)'' 
\newblock {\em Optimization Methods and Software}  {\bf 18} (2003)  491-505.

% cited
\bibitem{WAKI04}
H. Waki, S. Kim, M. Kojima, and M. Muramatsu,
``Sums of Squares and Semidefinite Programming Relaxations
for  Polynomial Optimization Problems
with Structured Sparsity'', 
% B-411, Dept. of Mathematical 
% and Computing Sciences, Tokyo Institute of 
% Technology, Oh-Okayama, Meguro, Tokyo 152-8552 Japan, October 2004. 
{\em SIAM J. Optimization} {\bf 17} (2006) 218--242.


\end{thebibliography}

 
