\thispagestyle{empty}

\noindent
\parbox[t]{1.2cm}{B-414}
\parbox[t]{14cm}
{User Manual for {\bf SparsePOP}:  
      a {\bf Sparse}  Semidefinite   Programming 
	Relaxation of {\bf P}polynomial {\bf O}ptimization {\bf P}roblems \\~\\
$\mbox{Hayato Waki}^{\star}$,
$\mbox{Sunyoung Kim}^{\dagger}$,
$\mbox{Masakazu Kojima}^{\ddagger}$, 
$\mbox{Masakazu Muramatsu}^{\sharp}$ \\
$\mbox{Hiroshi Sugimoto}^{\flat}$ \\
%\begin{center}
March 2005, revised December 2007
%\end{center}
} 

\vspace{1cm}

\noindent
{\bf Abstract.} \\ 
{\bf SparesPOP} is a MATLAB implementation of a sparse semidefinite programming (SDP)
relaxation method  for approximating a global optimal solution of a polynomial optimization problem (POP) 
proposed by Waki {\it et al}.
The sparse SDP relaxation exploits a sparse structure of polynomials in POPs
when applying ``a hierarchy of LMI relaxations of increasing dimensions''  by Lasserre.
The efficiency of {\bf SparsePOP} 
to approximate optimal solutions of POPs 
is thus increased, and  larger scale POPs can be handled. 
The software package {\bf SparesPOP}, this manual, and
a  test set of POPs from the literature are available at
\mbox{http://www.is.titech.ac.jp/$\sim$kojima/SparsePOP}.
%where the software package {\bf SparesPOP} and this manual can also be downloaded.

\vspace{1cm}
  
\noindent
{\bf Key words. } \vspace{0.1cm} \\
Polynomial optimization problem, sparsity, global optimization,
sums of squares optimization, 
semidefinite programming relaxation, MATLAB software package

\vspace{1cm}


\noindent
\parbox[t]{0.5cm}{$\star$}
\parbox[t]{14.9cm}{Department of Mathematical and Computing Sciences,
                      Tokyo Institute of Technology,
                      2-12-1 Oh-Okayama, Meguro-ku, Tokyo 152-8552 Japan.
{\it Hayato.Waki@is.titech.ac.jp}
}

\medskip

\noindent
\parbox[t]{0.5cm}{$\dagger$}
\parbox[t]{14.9cm}{Department of Mathematics, Ewha Women's University,
11-1 Dahyun-dong, Sudaemoon-gu, Seoul 120-750 Korea.
A considerable part of this work was conducted
while this author was visiting Tokyo Institute of Technology.
 The research was supported
by Kosef R01-2005-000-10271-0.
{\it skim@ewha.ac.kr}
}

\medskip

\noindent
\parbox[t]{0.5cm}{$\ddagger$}
\parbox[t]{14.9cm}{Department of Mathematical and Computing Sciences,
                      Tokyo Institute of Technology,
                      2-12-1 Oh-Okayama, Meguro-ku, Tokyo 152-8552 Japan.
                      Research supported by Grant-in-Aid for
                      Scientific Research on Priority Areas 16016234 and by Grant-in-Aid for
                      Scientific Research (B) 19310096.
{\it kojima@is.titech.ac.jp}
}


\medskip

\noindent
\parbox[t]{0.5cm}{$\sharp$}
\parbox[t]{14.9cm}{Department of  Computer Science, The University of
		Electro-Communications, Chofugaoka, Chofu-Shi, Tokyo
                      182-8585 Japan.
                      Research supported in part 
	by Grant-in-Aid for Young Scientists (B) 15740054.
                      {\it muramatu@cs.uec.ac.jp}
}

\medskip

\noindent
\parbox[t]{0.5cm}{$\flat$}
\parbox[t]{14.9cm}{Department of Mathematical and Computing Sciences,
                      Tokyo Institute of Technology,
                      2-12-1 Oh-Okayama, Meguro-ku, Tokyo 152-8552 Japan.
}

\newpage
